---
title: "foreign-body"
author: "Tyler Muffly, MD"
date: "8/16/2020"
output: html_document
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Set libPaths.
.libPaths("/Users/tylermuffly/.exploratory/R/4.0")

tinytex::install_tinytex()
# to uninstall TinyTeX, run tinytex::uninstall_tinytex()

# Load required packages.
library(janitor)
library(lubridate)
library(hms)
library(tidyr)
library(stringr)
library(readr)
library(forcats)
library(RcppRoll)
library(dplyr)
library(tibble)
library(bit64)
library(exploratory)
library(DiagrammeR)

# Steps to produce disposition
`disposition` <- exploratory::read_excel_file( "/Users/tylermuffly/Dropbox (Personal)/Trauma/disposition.xlsx", sheet = "Sheet1", na = c('','NA'), skip=0, col_names=TRUE, trim_ws=TRUE, tzone='America/Denver') %>%
  readr::type_convert() %>%
  exploratory::clean_data_frame() %>%
  mutate(disposition_description = `Treated and released, or examined and released without treatment 1`) %>%
  rename(disposition_code = `Treated and released, or examined and released without treatment 1`) %>%
  mutate_at(vars(everything()), funs(factor)) %>%
  mutate(disposition_code = word(disposition_code, -1, sep = "\\s+")) %>%
  mutate(disposition_description = recode(disposition_description, "Treated and transferred to another hospital 2" = "Treated and transferred to another hospital", "Treated and admitted for hospitalization (within same facility) 4" = "Treated and admitted for hospitalization (within same facility)", "Held for observation (includes admitted for observation) 5" = "Held for observation (includes admitted for observation)", "Left without being seen/Left against medical advice 6" = "Left without being seen/Left against medical advice", "Fatality, including DOA, died in the ED 8" = "Fatality, including DOA, died in the ED"))

# Steps to produce NEISSDiagnosisCodes
`NEISSDiagnosisCodes` <- exploratory::select_columns(exploratory::clean_data_frame(exploratory::read_delim_file("/Users/tylermuffly/Dropbox (Personal)/Trauma/NEISSDiagnosisCodes.csv" , ",", quote = "\"", skip = 0 , col_names = FALSE , na = c('','NA') , locale=readr::locale(encoding = "UTF-8", decimal_mark = ".", tz = "America/Denver", grouping_mark = "," ), trim_ws = TRUE , progress = FALSE)),"X1","X2") %>%
  readr::type_convert() %>%
  rename(Diagnosis = X1, Diagnosis_Code = X2) %>%
  mutate(Diagnosis = str_remove_all(Diagnosis, "[^a-zA-Z]+")) %>%
  mutate(Diagnosis = recode(Diagnosis, "DermatitisConjunctivitis" = "Dermatitis", "ElectricShock" = "Electric Shock", "ForeignBody" = "Foreign Body", "IngestedForeignObject" = "Ingested Foreign Object", "InternalOrganInjury" = "Internal Organ Injury", "NerveDamage" = "Nerve Damage", "StrainorSprain" = "Strain or Sprain", "SubmersionincludingDrowning" = "Submersion including Drowning", "OtherNotStated" = "Other")) %>%
  mutate_at(vars(everything()), funs(factor))

# Steps to produce products_NEISS
`products_NEISS` <- exploratory::read_delim_file("/Users/tylermuffly/Dropbox (Personal)/products_NEISS.csv" , ",", quote = "\"", skip = 0 , col_names = TRUE , na = c('','NA') , locale=readr::locale(encoding = "UTF-8", decimal_mark = ".", tz = "America/Denver", grouping_mark = "," ), trim_ws = TRUE , progress = FALSE) %>%
  readr::type_convert() %>%
  exploratory::clean_data_frame() %>%
  select(-X1) %>%
  mutate(code = factor(code)) %>%
  rename(products_description = title, products_code = code)

# Steps to produce Body_Part
`Body_Part` <- exploratory::read_excel_file( "/Users/tylermuffly/Dropbox (Personal)/My Mac (MacBook Pro)/Downloads/All sports injuries to pubic area/NEISS_FMT.XLSX", sheet = "NEISS_FMT", na = c('','NA'), skip=0, col_names=TRUE, trim_ws=TRUE, tzone='America/Denver') %>%
  readr::type_convert() %>%
  exploratory::clean_data_frame() %>%
  filter(`Format name` == "BDYPT") %>%
  select(-`Format name`, -`Ending value for format`) %>%
  mutate(`Format value label` = str_extract_all(`Format value label`, "[a-zA-Z]+")) %>%
  mutate(`Format value label` = list_to_text(`Format value label`), `Format value label` = str_to_title(`Format value label`)) %>%
  rename(BodyPartCode = `Starting value for format`, BodyPart_Text = `Format value label`) %>%
  mutate_at(vars(everything()), funs(factor))

# Steps to produce NEISS_2019_1
`NEISS_2019_1` <- exploratory::read_excel_file( "/Users/tylermuffly/Dropbox (Personal)/Trauma/NEISS_2019.XLSX", sheet = "NEISS_2019", na = c('','NA'), skip=0, col_names=TRUE, trim_ws=TRUE, tzone='America/Denver') %>%
  readr::type_convert() %>%
  exploratory::clean_data_frame()

# Steps to produce NEISS_2018_1
`NEISS_2018_1` <- exploratory::read_excel_file( "/Users/tylermuffly/Dropbox (Personal)/Trauma/NEISS_2018.XLSX", sheet = "NEISS_2018", na = c('','NA'), skip=0, col_names=TRUE, trim_ws=TRUE, tzone='America/Denver') %>%
  readr::type_convert() %>%
  exploratory::clean_data_frame()

# Steps to produce NEISS_2017_1
`NEISS_2017_1` <- exploratory::read_excel_file( "/Users/tylermuffly/Dropbox (Personal)/Trauma/NEISS_2017.XLSX", sheet = "NEISS_2017", na = c('','NA'), skip=0, col_names=TRUE, trim_ws=TRUE, tzone='America/Denver') %>%
  readr::type_convert() %>%
  exploratory::clean_data_frame()

# Steps to produce NEISS_2016_1
`NEISS_2016_1` <- exploratory::read_excel_file( "/Users/tylermuffly/Dropbox (Personal)/Trauma/NEISS_2016.XLSX", sheet = "NEISS_2016", na = c('','NA'), skip=0, col_names=TRUE, trim_ws=TRUE, tzone='America/Denver') %>%
  readr::type_convert() %>%
  exploratory::clean_data_frame()

# Steps to produce NEISS_2015_1
`NEISS_2015_1` <- exploratory::read_excel_file( "/Users/tylermuffly/Dropbox (Personal)/Trauma/NEISS_2015.XLSX", sheet = "NEISS_2015", na = c('','NA'), skip=0, col_names=TRUE, trim_ws=TRUE, tzone='America/Denver') %>%
  readr::type_convert() %>%
  exploratory::clean_data_frame()

# Steps to produce NEISS_2014_1
`NEISS_2014_1` <- exploratory::read_excel_file( "/Users/tylermuffly/Dropbox (Personal)/Trauma/NEISS_2014.XLSX", sheet = "NEISS_2014", na = c('','NA'), skip=0, col_names=TRUE, trim_ws=TRUE, tzone='America/Denver') %>%
  readr::type_convert() %>%
  exploratory::clean_data_frame()

# Steps to produce NEISS_2013_1
`NEISS_2013_1` <- exploratory::read_excel_file( "/Users/tylermuffly/Dropbox (Personal)/Trauma/NEISS_2013.XLSX", sheet = "NEISS_2013", na = c('','NA'), skip=0, col_names=TRUE, trim_ws=TRUE, tzone='America/Denver') %>%
  readr::type_convert() %>%
  exploratory::clean_data_frame()

# Steps to produce NEISS_2012_1
`NEISS_2012_1` <- exploratory::read_excel_file( "/Users/tylermuffly/Dropbox (Personal)/Trauma/NEISS_2012.XLSX", sheet = "NEISS_2012", na = c('','NA'), skip=0, col_names=TRUE, trim_ws=TRUE, tzone='America/Denver') %>%
  readr::type_convert() %>%
  exploratory::clean_data_frame()

# Steps to produce NEISS_2011_1
`NEISS_2011_1` <- exploratory::read_excel_file( "/Users/tylermuffly/Dropbox (Personal)/Trauma/NEISS_2011.XLSX", sheet = "NEISS_2011", na = c('','NA'), skip=0, col_names=TRUE, trim_ws=TRUE, tzone='America/Denver') %>%
  readr::type_convert() %>%
  exploratory::clean_data_frame()

# Steps to produce branching_point_1
`branching_point_1` <- 
  # All women from NEISS Years 2010 to 2019 with any diagnosis.  
  exploratory::read_excel_file( "/Users/tylermuffly/Dropbox (Personal)/Trauma/NEISS_2010.XLSX", sheet = "NEISS_2010", na = c('','NA'), skip=0, col_names=TRUE, trim_ws=TRUE, tzone='America/Denver') %>%
  readr::type_convert() %>%
  exploratory::clean_data_frame() %>%
  bind_rows(NEISS_2011_1, NEISS_2012_1, NEISS_2013_1, NEISS_2014_1, NEISS_2015_1, NEISS_2016_1, NEISS_2017_1, NEISS_2018_1, NEISS_2019_1, id_column_name = "ID", current_df_name = "NEISS_2010_1", force_data_type = TRUE) %>%
  select(-CPSC_Case_Number, -Sex, -`Hispanic Origin`, -Body_Part, -Other_Diagnosis, -Fire_Involvement) %>%
  mutate(Race = recode(Race, `2` = "Black", `0` = "NA", `1` = "White", `4` = "Asian", `3` = "Other", `5` = "NA", `6` = "NA")) %>%
  mutate(Race = na_if(Race, "NA")) %>%
  mutate_at(vars(Age, Race, Other_Race, Diagnosis, Body_Part_2, Diagnosis_2, Other_Diagnosis_2, Disposition, Location, `Alcohol involved`, `Drug involved`, Product_1, Product_2, Product_3, Stratum), funs(factor)) %>%
  left_join(Body_Part, by = c("Body_Part_2" = "BodyPartCode")) %>%
  select(-Body_Part_2) %>%
  left_join(products_NEISS, by = c("Product_1" = "products_code")) %>%
  left_join(products_NEISS, by = c("Product_2" = "products_code")) %>%
  left_join(products_NEISS, by = c("Product_3" = "products_code")) %>%
  select(-Product_1, -Product_2, -Product_3) %>%
  mutate(Age = parse_number(Age)) %>%
  filter(Age >= 16) %>%
  filter(Age < 100) %>%
  mutate(ID = word(ID, 2, sep = "\\s*\\_\\s*")) %>%
  mutate(ID = factor(ID)) %>%
  select(-Other_Race) %>%
  left_join(NEISSDiagnosisCodes, by = c("Diagnosis" = "Diagnosis_Code")) %>%
  left_join(NEISSDiagnosisCodes, by = c("Diagnosis_2" = "Diagnosis_Code")) %>%
  rename(Diagnosis1 = Diagnosis.y, Diagnosis2 = Diagnosis.y.y) %>%
  mutate(`Alcohol involved` = impute_na(`Alcohol involved`, type = "value", val = "0"), `Drug involved` = impute_na(`Drug involved`, type = "value", val = "0")) %>%
  left_join(disposition, by = c("Disposition" = "disposition_code")) %>%
  mutate(Treatment_Date_month = month(Treatment_Date, label = TRUE, abbr = FALSE), Treatment_Date_wday = wday(Treatment_Date, label = TRUE, abbr = FALSE)) %>%
  filter((is.na(Narrative) | Narrative != "47 YOF INJURED LIP WITH BB GUN LAST NIGHT. DX-SUPERIOR ORAL LABIAORAL FOREIGN BODY")) %>%
  mutate(Treatment_Date_month = recode(Treatment_Date_month, "March" = "Spring", "April" = "Spring", "May" = "Spring", "June" = "Summer", "July" = "Summer", "August" = "Summer", "September" = "Fall", "October" = "Fall", "November" = "Fall", "December" = "Winter", "January" = "Winter", "February" = "Winter")) %>%
  mutate(products_description.x = str_remove(products_description.x, regex(", not specified$", ignore_case = TRUE)), products_description.x = str_remove(products_description.x, regex(", not elsewhere classified", ignore_case = TRUE))) %>%
  mutate(products_description.x = str_remove(products_description.x, regex(", other or not specificed$", ignore_case = TRUE))) %>%
  mutate(products_description.x = str_to_title(products_description.x)) %>%
  select(-Diagnosis.x, -Diagnosis_2, -Other_Diagnosis_2, -Disposition, -Stratum, -PSU, -Weight)

# Steps to produce pregnant
`pregnant` <- `branching_point_1` %>%
  filter(str_detect(Narrative, regex("pregnan", ignore_case=TRUE)))

# Steps to produce rectum
`rectum` <- `branching_point_1` %>%
  filter(str_detect(Narrative, regex("rectum", ignore_case=TRUE)))

# Steps to produce bladder
`bladder` <- `branching_point_1` %>%
  filter(str_detect(Narrative, regex("bladder", ignore_case=TRUE))) %>%
  mutate(calculation_1 = Treatment_Date_wday) %>%
  mutate(calculation_1 = recode(calculation_1, "Tuesday" = "weekday", "Monday" = "weekday", "Wednesday" = "weekday", "Sunday" = "weekend", "Friday" = "weekday", "Saturday" = "weekend", "Thursday" = "weekday"))

# Steps to produce clitoris
`clitoris` <- `branching_point_1` %>%
  filter(str_detect(Narrative, regex("clitor", ignore_case=TRUE)))

# Steps to produce vagina
`vagina` <- `branching_point_1` %>%
  filter(str_detect(Narrative, regex("vagina", ignore_case=TRUE)))

# Steps to produce uterus
`uterus` <- `branching_point_1` %>%
  filter(str_detect(Narrative, regex("uterus", ignore_case=TRUE)))

# Steps to produce perineum
`perineum` <- `branching_point_1` %>%
  filter(str_detect(Narrative, regex("perineum", ignore_case=TRUE)))

# Steps to produce ovary
`ovary` <- `branching_point_1` %>%
  filter(str_detect(Narrative, regex("ovar", ignore_case=TRUE)))

# Steps to produce the output
df <- `branching_point_1` %>%
 
  # We were able to search for injuries with the anatomic designations including “bladder,” “clitoris,” “labia,” “ovary,” "perineum,"  "rectum," "uterus," and “vagina. ” We excluded pregnant women with injuries in this analysis.
  filter(str_detect(Narrative, regex("labia", ignore_case=TRUE))) %>%
  bind_rows(ovary, perineum, uterus, vagina, clitoris, bladder, rectum, pregnant, id_column_name = "ID", current_df_name = "everyone_labia", force_data_type = TRUE) %>%
  
  # Remove pregnant women.  Instead of relying on the NEISS data for categories I used word recognition within the narratives for all females over the decade of interest.  
  filter((is.na(ID.new) | ID.new != "pregnant")) %>%
  
  # I kept only unique narrations to avoid double counting one person.  
  distinct(Narrative, .keep_all = TRUE) %>%
  mutate_at(vars(ID.new, ID, Race, `Alcohol involved`, `Drug involved`, BodyPart_Text, products_description.x, products_description.y, products_description, Diagnosis1, Diagnosis2, disposition_description, Treatment_Date_month, Treatment_Date_wday), funs(factor)) %>%
  
  # R sort of freaks out when column names are not a single string of characters.  Therefore I put the underscore between Anatomy and injured.  
  rename(Anatomy_injured = ID.new) %>%
  rename(Year_of_injury = ID, products_description1 = products_description.x, products_description2 = products_description.y, products_description3 = products_description) %>%
  
  # Merging variable of shaver together.  Could be the next paper:  Shaving injuries in the ED?
  mutate(products_description1 = recode(products_description1, "nonelectric razors or shavers" = "razors or shavers", "razors or shavers, not specified" = "razors or shavers", "electric razors or shavers" = "razors or shavers")) %>%
  mutate(Age = as.integer(Age)) %>%
  
  # I am limiting the anatomy sites to vagina, rectum and bladder.  
  filter(Anatomy_injured %in% c("vagina", "rectum", "bladder")) %>%
  
  # Are people more likely to come in with a foreign body on a particular day of the week?
  mutate(Treatment_Date_wday = fct_relevel(Treatment_Date_wday, "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")) %>%
  select(-BodyPart_Text, -Diagnosis2) %>%
  
  # Are people more likely to come in with a foreign body on a holiday?
  mutate(Holiday = tis::isHoliday(Treatment_Date, goodFriday = FALSE, board = TRUE), Holiday = factor(Holiday)) %>%
  mutate_at(vars(Treatment_Date_month, Anatomy_injured), funs(factor)) %>%
  mutate(Age = as.numeric(Age)) %>%
  filter((is.na(Anatomy_injured) | Anatomy_injured != "bladder")) %>%
  mutate(Anatomy_injured = fct_relevel(Anatomy_injured, "rectum", "vagina"), Anatomy_injured = fct_other(Anatomy_injured, drop = c("bladder")), Anatomy_injured = fct_drop(Anatomy_injured), Anatomy_injured = as.character(Anatomy_injured)) %>%
  mutate(Age_category = cut(Age, breaks = 10, dig.lab = 10)) %>%
  filter(Diagnosis1 == "Foreign Body") %>%
  select(-calculation_1)
```

There are `r nrow(df)` women over age 16 between 2010 to 2020 that had a retained foreign body in the vagina, bladder, or rectum.  

```{r}
colnames(df)

## Table 1

all_data_ordered <- df %>%
  exploratory::reorder_cols(c("Age", "Race", "Year_of_injury", "Anatomy_injured", "products_description1", "disposition_description", "Treatment_Date_month"))

Table_1 <- arsenal::tableby(formula = ~
                              Age + 
                              Race + 
                              Anatomy_injured  + 
                              #products_description1  + 
                              disposition_description +  
                              Treatment_Date_month, 
                            
                            data=all_data_ordered, 
                            control =arsenal::tableby.control(test = TRUE,
                                                              total = TRUE,
                                                              digits = 1L, digits.count = 0L, cat.simplify = FALSE, numeric.simplify = TRUE,numeric.stats = c("Nmiss2", "median",  "q1q3"), cat.stats =c("Nmiss2", "countpct"), stats.labels = list(Nmiss = "N Missing", Nmiss2 ="N Missing", meansd = "Mean (SD)", medianrange = "Median (Range)",median ="Median", medianq1q3 = "Median (Q1, Q3)",q1q3 = "Q1, Q3",iqr = "IQR",range = "Range",countpct = "Count (Pct)", Nevents = "Events", medSurv ="Median Survival",medTime = "Median Follow-Up")))

summary(Table_1,
        text=T,
        title = 'Table: Patient Descriptive Variables of Females Presenting to the ED with Pelvic Foreign Body from 2010 to 2020', 
        #labelTranslations = mylabels, #Seen in additional functions file
        pfootnote=TRUE)
```


```{r}
#formatC(1:10 * 100000, format="d", big.mark=",")

nationalestimate  <- 540860 #Female foreign bodies per year
femaleforeignbodyfor2019 <- 2758  #all anatomic sites
femaleforeignbodyfordecade <- femaleforeignbodyfor2019 * 10 #yearly cases times 10 years of cases
percentagefemalepelvicforeignbody <- round(femaleforeignbodyfordecade/nationalestimate,4) * 100
```

```{r}
tab1 <- prop.table(table(df$disposition_description))*100
tab1[[4]]
```

```{r}
tab2 <- prop.table(table(df$Anatomy_injured))*100
tab2

#https://stackoverflow.com/questions/17374651/find-the-n-most-common-values-in-a-vector/17374663#17374663
```

```{r}
tab3 <- prop.table(table(df$Locale_Description))*100
tab3
```

```{r}
#Holiday
tab4 <- prop.table(table(df$Holiday))*100
tab4
```
```{r}
tab5 <- df %>% summarize_group(group_cols = c("Anatomy_injured" = "Anatomy_injured"),group_funs = c("none"),Age_med = median(Age, na.rm = TRUE))
```

## Materials & Methods
```r
#https://bookdown.org/yihui/rmarkdown-cookbook/diagrams.html

DiagrammeR::grViz("digraph {
  graph [layout = dot, rankdir = TB]
  
  node [shape = rectangle]        
  rec1 [label = 'Step 1. Download 2010 through 2019 data on all women from \nNational Electronic Injury Surveillance System (NEISS) \n N= 1,107,712']
  rec2 [label = 'Step 2. Filter Patient Narrative for terms Rectum and Vagina \nN = 5,752']
  rec3 [label =  'Step 3. Exclude Patients who were Pregnant \nN = 2,550']
  rec4 [label = 'Step 4. Include Patients who were Diagnosed with Foreign Body \nN= 436']
  
  # edge definitions with the node IDs
  rec1 -> rec2 -> rec3 -> rec4
  }", 
  height = 500)
```

## Results
Over the 10‐year period from `r format(min(df$Treatment_Date), format="%B %d %Y")` to `r format(max(df$Treatment_Date), format="%B %d %Y")`, there were an estimated `r formatC(nationalestimate, format = "d", big.mark=",")` ED visits for retained foreign body, and `r formatC(femaleforeignbodyfordecade, format = "d", big.mark = ",")` (`r percentagefemalepelvicforeignbody`%) ED visits were related to female pelvic injuries as estimated using our sample size of `r nrow(df)` patients. The typical patient with a retained bladder, rectal, or vaginal foreign body was a `r median(df$Age)`year-old (IQR: `r median(df$Age, na.rm=TRUE)-IQR(df$Age, na.rm=TRUE)` - `r IQR(df$Age, na.rm=TRUE)+median(df$Age, na.rm=TRUE)`))  `r names(which.max(table(df$Race)))` female who stated that they were not using drugs or alcohol during the placement of the foreign body.  The most common foreign body of the `r english::as.english(nlevels(df$products_description1))` types of foreign body found in the vagina and rectum was `r tolower(names(which.max(table(df$products_description1))))`, followed by `r tolower(names(summary(as.factor(df$products_description1))))[[2]]` and `r tolower(names(summary(as.factor(df$products_description1))))[[3]]`. Many of the vibrators were lost during sexual activity.  Jewelry mainly consisted of infected vaginal piercings that were removed for abscess drainage.  Drugs were commonly confiscated from the vagina following an arrest. The caps of writing instruments, perfume, hairspray, enema instruments, and toothpaste were parts that became foreign bodies.  Most injuries presented to the ED on a `r names(which.max(table(df$Treatment_Date_wday)))` during the `r tolower(names(which.max(table(df$Treatment_Date_month))))` months, and approximately `r round(tab4[2], 1)`% of foreign body cases presented on a federal holiday.  

Of the `r nrow(df)` visits for foreign body in the female pelvis that were analyzed, 
(`r round(tab1[[4]],1)`%) visits involved patients who were admitted to the hospital; `r round(tab1[[3]],1)`% left the ED against medical advice; and `r round(tab1[[2]],1)`% were held for observation. Overall (Fig. 1) incidence of injury and anatomic site‐specific (Fig. 2) incidence of injury fluctuated during the time period studied. Upon analysis of foreign body by anatomic site, a majority of visits (`r round(tab2[[2]],1)`%) was due to injuries to the `r names(which.max(table(df$Anatomy_injured)))`. A plurality of vaginal foreign body patients had a vibrator as a vaginal foreign body specifically vibrating eggs, batteries that fell out of a vibrator, and penis rings.   The `r names(which.min(table(df$Anatomy_injured)))` was the least common sites of injury (Fig. 3).  Patients with a rectal foreign body were mainly anal plugs, vibrators confirmed by x-rays, and beads.  No patient presented with both a rectal and vaginal foreign body present at the same time.  Patients with vaginal foreign body were significantly younger than those with rectal injury (`r tab5[[1,2]]` vs `r tab5[[2,2]]`, .  





